---
title: "Project 3: Proj3Package Tutorial"
author: "Todd Meng"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Proj3Package Tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Installing Package

This package contains useful statistical functions for t-tests, linear regression, K nearest neighbors cross-validation, and random forest cross-validation. It also contains the data my_gapminder. To install from github, use the devtools package.

```{r, eval=FALSE}
library(devtools)
# basic install
install_github("toadmengo/Proj3Package")
# install with vignette
install_github("toadmengo/Proj3Package", 
               build_vignette = TRUE, build_opts = c())
```

```{r setup}
library(Proj3Package)
```

### my_t.test

my_t.test runs a t-test on data. It has parameters: x is the numeric data, alternative is a character input that signifies the type of t-test, and mu is the null hypothesis value. alternative must be either "greater", "less", or "two.sided". The following demonstrates a two-sided t-test on the my_gapminder life-expectancy data. It tests the hypothesis that mu != 60.

```{r}
x <- my_gapminder$lifeExp
# two sided test
my_t.test(x, "two.sided", 60)
```

We can test for mu < 60.

```{r}
# one-sided less-than test
my_t.test(x, "less", 60)
```

We can test mu >60.
```{r}
# one-sided less-than test
my_t.test(x, "greater", 60)
```
